Discrepancy, Coresets, and Sketches in Machine Learning

### 2019-01-29 komlos.tex
switch section order

\begin{abstract}
This paper defines the notion of class discrepancy for families of functions.  It shows that low discrepancy classes admit small offline and streaming coresets. We provide general techniques for bounding the discrepancy of machine learning problems and specifically do so for matrix covariance approximation, logistic regression, kernel density and any analytic function of the dot product or the squared distance.
Our result resolves a long standing open question regarding the coreset complexity of Gaussian kernel density estimation.  
We provide two more related but independent results. 
First, an exponential improvement of the widely used merge-and-reduce trick which gives improved streaming sketches for any low discrepancy problem.
Second, an extremely simple algorithm for finding low discrepancy sequences (and therefore coresets) for any positive semi-definite kernel. 
This paper makes an explicit connection between low discrepancy, coresets, learnability, and streaming algorithms. 
\end{abstract}

\section{Introduction}
In both machine learning and in streaming and sketching problems our goal is (often) to approximate sums or expectations of well behaved functions.
Specifically, we need to approximate $\E_x f(x)$ or $\frac{1}{n}\sum_{i=1}^{n} f(x_i)$ for every $f\in \F$ where $\F$ is a family of functions and $x_i \in \X$ are either sampled training examples or an arbitrary set of stream items. 
Standard generalization results show that for a large enough value of $n$ the average approximates the mean if the complexity of $\F$ is bounded and the samples $x_i$ are drawn i.i.d.\ from an underlying distribution. We therefore focus on approximating the average, or rather, the sum $\sum_{i=1}^{n} f(x_i)$. For notational convenience we use a parameter $q \in \Q$ to index into $\F$ explicitly. 
In other words, there is a bijective mapping between $\Q \equiv \F$ such that $f(x) \in \F$ iff there exists $q\in\Q$ such that $f(x,q)=f(x)$.
We keep using the two different functions $f:\X\rightarrow\R$ and $f:\X,\Q\rightarrow\R$ interchangeably. 
One should think about $q$ as either the model parameters or a query for the sketch.

@@

\todo{Write short intro for this section}
\subsection{Class Discrepancy}
We begin by giving three equivalent definitions of complexity based on discrepancy for sets, functions, and function families. We will use all three interchangeably throughout the manuscript.

\begin{definition}[\bf Class Discrepancy] \label{cd1}
Let $A \subset \R^m$ and $\sigma \in \{-1,1\}^m$ the class discrepancy of $A$ is 
\[
K_m(A)  = \min_\sigma \max_{a\in A} \left| \frac{1}{m} \sum_{i=1}^{m}  \sigma_i  a_i\right| 
\]
\end{definition}
%
\begin{definition}[\bf Class Discrepancy]\label{cd2}
Let $f:\X,\Q\rightarrow\R$ and $\sigma \in \{-1,1\}^m$. 
The class discrepancy of $f$ w.r.t.\ $\{x_1,\ldots,x_m\} \subset \X$ is 
\[
K_m(f) =  \min_{\sigma} \max_{q \in \Q}  \left|\frac{1}{m} \sum_{i=1}^{m}\sigma_i f(x_i,q)\right|
\]
\end{definition}
%
\begin{definition}[\bf Class Discrepancy]\label{cd3}
Let $\F$ be a family of functions $f:\X\rightarrow\R$ and $\sigma \in \{-1,1\}^m$. 
The class discrepancy of $\F$ w.r.t.\ $\{x_1,\ldots,x_m\} \subset \X$ is 
\[
K_m(\F) =  \min_{\sigma} \max_{f \in \F}  \left| \frac{1}{m} \sum_{i=1}^{m}\sigma_i f(x_i)\right|
\]
\end{definition}
The class discrepancy of $f$ or $\F$ without a reference a set $\{x_1,\ldots,x_m\}$ is the upper bound on any subset of $\X$ of size $m$.
Throughout the manuscript we assume a bijective mapping between $\F$ and $\Q$. 
Specifically, $f \in \F$ if there exists $q \in \Q$ such that $f(x) = f(x,q)$ and vice versa. 
In the context of machine learning, one should think about $f(x,q)$ as the loss associated with example $x$ and model parameters $q$.
Definitions~\ref{cd2} and \ref{cd3} are therefore equivalent. 
For the equivalence of Definition~\ref{cd1}, the set $A$ should be thought of as the set of possible induces losses. 
Namely, $a\in A$ if there is a model $q$ such $a_i = f(x_i,q)$.
The three equivalent definitions between the definitions is therefore straight forward.

@@

\section{Related Work}
\begin{itemize}
\item coresets - motivation, some papers, streaming algorithms for coresets.

>>>>

\begin{abstract}
This paper defines the notion of class discrepancy for families of functions.  It shows that low discrepancy classes admit small offline and streaming coresets. We provide general techniques for bounding the discrepancy of machine learning problems and specifically do so for matrix covariance approximation, logistic regression, kernel density and any analytic function of the dot product or the squared distance.
Our result resolves a long standing open problem regarding the coreset complexity of Gaussian kernel density estimation.  
We provide two more related but independent results. 
First, an exponential improvement of the widely used merge-and-reduce trick which gives improved streaming sketches for any low discrepancy problem.
Second, an extremely simple algorithm for finding low discrepancy sequences (and therefore coresets) for any positive semi-definite kernel. 
This paper establishes some explicit connections between low class discrepancy, coreset complexity, learnability, and streaming algorithms. 
\end{abstract}

\section{Introduction}

\begin{itemize}
\item coresets - motivation, some papers, streaming algorithms for coresets.

In both machine learning and in streaming and sketching problems our goal is (often) to approximate sums or expectations of well behaved functions.
Specifically, we need to approximate $\E_x f(x)$ or $\frac{1}{n}\sum_{i=1}^{n} f(x_i)$ for every $f\in \F$ where $\F$ is a family of functions and $x_i \in \X$ are either sampled training examples or an arbitrary set of stream items. 
Standard generalization results show that for a large enough value of $n$ the average approximates the mean if the complexity of $\F$ is bounded and the samples $x_i$ are drawn i.i.d.\ from an underlying distribution. We therefore focus on approximating the average, or rather, the sum $\sum_{i=1}^{n} f(x_i)$. For notational convenience we use a parameter $q \in \Q$ to index into $\F$ explicitly. 
In other words, there is a bijective mapping between $\Q \equiv \F$ such that $f(x) \in \F$ iff there exists $q\in\Q$ such that $f(x,q)=f(x)$.
We keep using the two different functions $f:\X\rightarrow\R$ and $f:\X,\Q\rightarrow\R$ interchangeably. 
One should think about $q$ as either the model parameters or a query for the sketch.

@@

\subsection{Class Discrepancy}
We begin by giving three equivalent definitions of complexity based on discrepancy for sets, functions, and function families. We will use all three interchangeably throughout the manuscript.

\begin{definition}[\bf Class Discrepancy] \label{cd1}
Let $A \subset \R^m$ and $\sigma \in \{-1,1\}^m$ the class discrepancy of $A$ is 
$
K_m(A)  = \min_\sigma \max_{a\in A} \left| \frac{1}{m} \sum_{i=1}^{m}  \sigma_i  a_i\right| 
$.
\end{definition}
%
\begin{definition}[\bf Class Discrepancy]\label{cd2}
Let $f:\X,\Q\rightarrow\R$ and $\sigma \in \{-1,1\}^m$. 
The class discrepancy of $f$ w.r.t.\ $\{x_1,\ldots,x_m\} \subset \X$ is 
$
K_m(f) =  \min_{\sigma} \max_{q \in \Q}  \left|\frac{1}{m} \sum_{i=1}^{m}\sigma_i f(x_i,q)\right|
$.
\end{definition}
%
\begin{definition}[\bf Class Discrepancy]\label{cd3}
Let $\F$ be a family of functions $f:\X\rightarrow\R$ and $\sigma \in \{-1,1\}^m$. 
The class discrepancy of $\F$ w.r.t.\ $\{x_1,\ldots,x_m\} \subset \X$ is 
$
K_m(\F) =  \min_{\sigma} \max_{f \in \F}  \left| \frac{1}{m} \sum_{i=1}^{m}\sigma_i f(x_i)\right|
$.
\end{definition}
The class discrepancy of $f$ or $\F$ without a reference a set $\{x_1,\ldots,x_m\}$ is the upper bound on any subset of $\X$ of size $m$.
Throughout the manuscript we assume a bijective mapping between $\F$ and $\Q$. 
Specifically, $f \in \F$ if there exists $q \in \Q$ such that $f(x) = f(x,q)$ and vice versa. 
In the context of machine learning, one should think about $f(x,q)$ as the loss associated with example $x$ and model parameters $q$.
Definitions~\ref{cd2} and \ref{cd3} are therefore equivalent. 
For the equivalence of Definition~\ref{cd1}, the set $A$ should be thought of as the set of possible induces losses. 
Namely, $a\in A$ if there is a model $q$ such $a_i = f(x_i,q)$.
The three equivalent definitions between the definitions is therefore straight forward.

@@
